Modeling the uncertainties of solar-system ephemerides for robust gravitational-wave searches with pulsar timing arrays

### 2020-01-22 paper-apj/main.tex
Submitted ApJ version

\documentclass[iop,apj,twocolappendix]{emulateapj}

%\documentclass[reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
%amsmath,amssymb,
%aps,prd,floatfix,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
%]{revtex4-1}

\usepackage{natbib}
\usepackage{amsmath,amssymb}
\usepackage{apjfonts}
%\usepackage{caption}
%\usepackage{url}

\usepackage[usenames]{xcolor}
% matplotlib colors for text use
\definecolor{mplblue}{HTML}{1F77B4}
\definecolor{mplorange}{HTML}{FF7F0E}
\definecolor{mplgreen}{HTML}{2CA02C}
\definecolor{mplred}{HTML}{D62728}

\usepackage[breaklinks,plainpages=false,colorlinks=true,anchorcolor=cyan,linkcolor=mplred,citecolor=cyan,urlcolor=magenta]{hyperref}
% \usepackage[backref, breaklinks, plainpages=false,colorlinks=true, anchorcolor=cyan, linkcolor=mplred, citecolor=cyan, urlcolor=magenta, bookmarks=false]{hyperref}
% \usepackage[all]{hypcap}
% \usepackage[caption=false]{subfig}
% \renewcommand*{\backref}[1]{[#1]}

\citestyle{apj}

@@

% \usepackage{graphicx}% Include figure files
% \usepackage{dcolumn}% Align table columns on decimal point
% \usepackage{bm}% bold math
%\usepackage[backref, breaklinks, plainpages=false, colorlinks=true, anchorcolor=cyan, linkcolor=red, citecolor=cyan, filecolor=magenta, urlcolor=magenta]{hyperref}
% \usepackage[breaklinks, plainpages=false, colorlinks=true, anchorcolor=cyan, linkcolor=red, citecolor=cyan, filecolor=magenta, urlcolor=magenta]{hyperref}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
% \usepackage{xcolor}
% \usepackage{wasysym}
% \usepackage{enumitem}

% \newcommand{\srt}[1]{{\color{magenta}[SRT: #1]}}
% \newcommand{\mv}[1]{{\color{red}[MV: #1]}}
% \newcommand{\jjs}[1]{{\color{blue}[JJS: #1]}}

@@

%\preprint{APS/123-QED}

\title{Modeling the uncertainties of solar-system ephemerides for robust gravitational-wave searches with pulsar timing arrays}
%\thanks{A footnote to the article title}

\input{authors}

% named authors: M.~Vallisneri, S.~R.~Taylor, J.~Simon, W.~M.~Folkner, R.~S.~Park, C.~Cutler, J.~A.~Ellis, T.~J.~W.~Lazio, S.~J.~Vigeland
% \email{Second.Author@institution.edu}
% \affiliation{California Institute of Technology, Pasadena CA 91125}%

% \collaboration{NANOGrav Collaboration}%\noaffiliation

% \date{\today}

\begin{abstract}
The regularity of pulsar emissions becomes apparent once we reference the pulses' times of arrivals to the inertial rest frame of the solar system. It follows that errors in the determination of Earth's position with respect to the solar-system barycenter can appear as a time-correlated bias in pulsar-timing residual time series, affecting the searches for low-frequency gravitational waves performed with pulsar timing arrays.
Indeed, recent array datasets yield different gravitational-wave background upper limits and detection statistics when analyzed with different solar-system ephemerides. Crucially, the ephemerides do not generally provide usable error representations.
In this article we describe the motivation, construction, and application of a physical model of solar-system ephemeris uncertainties, which focuses on the degrees of freedom (Jupiter's orbital elements) most relevant to gravitational-wave searches with pulsar timing arrays.
This model, \textsc{BayesEphem}, was used to derive ephemeris-robust results in NANOGrav's 11-yr stochastic-background search, and it provides a foundation for future searches by NANOGrav and other consortia.
The analysis and simulations reported here suggest that ephemeris modeling reduces the gravitational-wave sensitivity of the 11-yr dataset; and that this degeneracy will vanish with improved ephemerides and with the longer pulsar timing datasets that will become available in the near future.
\end{abstract}

%\pacs{Valid PACS appear here}
%\keywords{Suggested keywords}

\maketitle

%\tableofcontents

@@

This article is organized as follows. In Sec.\ \ref{sec:sses} we summarize the SSE production, and we describe the history and stated accuracy of the JPL SSEs adopted in our work; in Sec.\ \ref{sec:systematics} we discuss the TOA delays induced by SSE errors, which are partially absorbed by the timing model, and we identify Jupiter and Saturn's orbits as the drivers behind GW-posterior discrepancies; in Sec.\ \ref{sec:physical} we formulate \textsc{BayesEphem}, and give details about its implementation in our PTA data-analysis software, \textsc{Enterprise} \citep{enterprise}; in Sec.\ \ref{sec:results} we report on the Bayesian posteriors (for the GW amplitude and for orbital-correction parameters) obtained with \textsc{BayesEphem} for NANOGrav's 11-yr dataset, reproducing and expanding the results of \cite{2018ApJ...859...47A};
in Sec.\ \ref{sec:simulations} we present simulations that probe the reduction in GW sensitivity due to \textsc{BayesEphem};
in Sec.\ \ref{sec:othermodels} we discuss other approaches toward SSE uncertainty modeling;
last, in Sec.\ \ref{sec:conclusion} we offer our brief conclusions.

All computational results presented in this article were obtained with \textsc{Enterprise}  \citep{enterprise}, used in conjunction with the stochastic sampler \textsc{PTMCMCSampler}\footnote{\href{https://github.com/jellis18/PTMCMCSampler}{https://github.com/jellis18/PTMCMCSampler}} and the pulsar-timing package \textsc{Tempo2} \citep{2012ascl.soft10015H}.

@@

These corrections yield delays with the same periods as the corresponding planets, so again they may be observable in PTA datasets for Jupiter and Saturn, but not for Uranus and Neptune.
% MV: say anything about the decay of accuracy past the end of the data?
%
\begin{figure*}[t]
    \centering
    \includegraphics[width=2\columnwidth]{figures/roemerdiff.pdf}
    \caption{Differences in R{\o}mer delays computed with DE421/DE430/DE435/DE438 and taking DE436 as reference for 100 simulated pulsars randomly distributed across the sky, plotted after subtracting the best-fitting quadratic and yearly sinusoid. The darker curves with various dashing styles allow a comparison of the same pulsars across the three differences. The plot covers the span of the NANOGrav 11-yr dataset.
    For any given pulsar, these differences are reduced to less than 10 ns after subtracting the best-fitting linear combination of \textsc{BayesEphem} partials, dotted into the pulsar position to obtain R{\o}mer delays.
    }
    \label{fig:Roemer}
\end{figure*}

@@

\emph{Acknowledgments.} \input{acknowledge}

@@

% \newpage
\bibliographystyle{yahapj}
\bibliography{references}

>>>>

\documentclass{aastex63}

\usepackage{amsmath}
\usepackage{amsfonts}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex63}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design 
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%   tighten      : Makes baselineskip slightly smaller, only works with 
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for 
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much 
%%                  longer author list but may be desirable for short 
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex63}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{June 1, 2019}
%\revised{January 10, 2019}
%\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the argument.
\submitjournal{Astrophys. J.}

%% For manuscript that include authors in collaborations, AASTeX v6.3
%% builds on the \collaboration command to allow greater freedom to 
%% keep the traditional author+affiliation information but only show
%% subsets. The \collaboration command now must appear AFTER the group
%% of authors in the collaboration and it takes TWO arguments. The last
%% is still the collaboration identifier. The text given in this
%% argument is what will be shown in the manuscript. The first argument
%% is the number of author above the \collaboration command to show with
%% the collaboration text. If there are authors that are not part of any
%% collaboration the \nocollaboration command is used. This command takes
%% one argument which is also the number of authors above to show. A
%% dashed line is shown to indicate no collaboration. This example manuscript
%% shows how these commands work to display specific set of authors 
%% on the front page.
%%
%% For manuscript without any need to use \collaboration the 
%% \AuthorCollaborationLimit command from v6.2 can still be used to 
%% show a subset of authors.
%
%\AuthorCollaborationLimit=2
%
%% will only show Schwarz & Muench on the front page of the manuscript
%% (assuming the \collaboration and \nocollaboration commands are
%% commented out).
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

@@

% \newcommand{\eqref}[1]{(\ref{#1}}

\graphicspath{{./}{figures/}}

@@

\title{Modeling the uncertainties of solar-system ephemerides for robust gravitational-wave searches with pulsar timing arrays}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.3 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affiliations. The change was necessary to implement 
%% auto-indexing of affiliations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% argument which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
%% present or in the LaTeX pathway, the OrcID icon will appear next to
%% the authors name.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.3 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.3 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Michele Vallisneri}
\email{Michele.Vallisneri@jpl.nasa.gov}

\author[0000-0002-4162-0033]{M. Vallisneri}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\author[0000-0003-0264-1453]{S. R. Taylor}
\affiliation{Department of Physics and Astronomy, Vanderbilt University, 2301 Vanderbilt Place, Nashville, TN 37235, USA}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\author[0000-0003-1407-6607]{J. Simon}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\author[0000-0001-5133-9934]{W. M. Folkner}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\author[0000-0001-9896-4585]{R. S. Park}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\author[0000-0002-2080-1468]{C. Cutler}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\author[0000-0003-1715-974X]{J. A. Ellis}
\affiliation{Infinia ML, 202 Rigsbee Avenue, Durham, NC 27701, USA}
\author{T. J. W. Lazio}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\author[0000-0003-4700-9072]{S. J. Vigeland}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\author[0000-0002-2059-0525]{K. Aggarwal}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author{Z. Arzoumanian}
\affiliation{X-Ray Astrophysics Laboratory, NASA Goddard Space Flight Center, Code 662, Greenbelt, MD 20771, USA}
\author[0000-0003-2745-753X]{P. T. Baker}
\affiliation{Department of Physics and Astronomy, Widener University, One University Place, Chester, PA 19013, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0001-6341-7178]{A. Brazier}
\affiliation{Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}
\affiliation{Cornell Center for Advanced Computing, Ithaca, NY 14853, USA}
\author[0000-0003-3053-6538]{P. R. Brook}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0003-4052-7838]{S. Burke-Spolaor}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0002-2878-1502]{S. Chatterjee}
\affiliation{Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}
\author[0000-0002-4049-1882]{J. M. Cordes}
\affiliation{Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}
\author[0000-0002-7435-0869]{N. J. Cornish}
\affiliation{Department of Physics, Montana State University, Bozeman, MT 59717, USA}
\author[0000-0002-2578-0360]{F. Crawford}
\affiliation{Department of Physics and Astronomy, Franklin \& Marshall College, P.O.~Box 3003, Lancaster, PA 17604, USA}
\author[0000-0002-6039-692X]{H. T. Cromartie}
\affiliation{University of Virginia, Department of Astronomy, P.O.~Box 400325, Charlottesville, VA 22904, USA}
\author[0000-0002-1529-5169]{K. Crowter}
\affiliation{Department of Physics and Astronomy, University of British Columbia, 6224 Agricultural Road, Vancouver, BC V6T 1Z1, Canada}
\author[0000-0002-2185-1790]{M. DeCesar}
\affiliation{Department of Physics, Lafayette College, Easton, PA 18042, USA}
\author[0000-0002-6664-965X]{P. B. Demorest}
\affiliation{National Radio Astronomy Observatory, 1003 Lopezville Rd., Socorro, NM 87801, USA}
\author[0000-0001-8885-6388]{T. Dolch}
\affiliation{Department of Physics, Hillsdale College, 33 E.~College Street, Hillsdale, MI 49242, USA}
\author[0000-0002-2223-1235]{R. D. Ferdman}
\affiliation{Department of Physics, University of East Anglia, Norwich, UK}
\author[0000-0001-7828-7708]{E. C. Ferrara}
\affiliation{NASA Goddard Space Flight Center, Greenbelt, MD 20771, USA}
\author[0000-0001-8384-5049]{E. Fonseca}
\affiliation{Department of Physics, McGill University, 3600  University St., Montreal, QC H3A 2T8, Canada}
\author[0000-0001-6166-9646]{N. Garver-Daniels}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0001-8158-683X]{P. Gentile}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0003-1884-348X]{D. Good}
\affiliation{Department of Physics and Astronomy, University of British Columbia, 6224 Agricultural Road, Vancouver, BC V6T 1Z1, Canada}
\author[0000-0003-2742-3321]{J. S. Hazboun}
\affiliation{University of Washington Bothell, 18115 Campus Way NE, Bothell, WA 98011, USA}
\author[0000-0003-4143-8132]{A. M. Holgado}
\affiliation{NCSA and Department of Astronomy, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA}
\author[0000-0002-9682-3604]{E. A. Huerta}
\affiliation{NCSA and Department of Astronomy, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA}
\author[0000-0001-5123-9917]{K. Islo}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\author[0000-0003-1082-2342]{R. Jennings}
\affiliation{Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}
\author{G. Jones}
\affiliation{Department of Physics, Columbia University, New York, NY 10027, USA}
\author[0000-0001-6607-3710]{M. L. Jones}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\author[0000-0001-6295-2881]{D. L. Kaplan}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\author[0000-0002-6625-6450]{L. Z. Kelley}
\affiliation{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), Northwestern University, Evanston, IL 60208}
\author[0000-0003-0123-7600]{J. S. Key}
\affiliation{University of Washington Bothell, 18115 Campus Way NE, Bothell, WA 98011, USA}
\author[0000-0003-0721-651X]{M. T. Lam}
\affiliation{School of Physics and Astronomy, Rochester Institute of Technology, Rochester, NY 14623, USA}
\affiliation{Laboratory for Multiwavelength Astronomy, Rochester Institute of Technology, Rochester, NY 14623, USA}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0002-2034-2986]{L. Levin}
\affiliation{Jodrell Bank Centre for Astrophysics, University of Manchester, Manchester, M13 9PL, United Kingdom}
\author[0000-0003-1301-966X]{D. R. Lorimer}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0001-5373-5914]{J. Luo}
\affiliation{Department of Physics, McGill University, 3600  University St., Montreal, QC H3A 2T8, Canada}
\author[0000-0001-5229-7430]{R. S. Lynch}
\affiliation{Green Bank Observatory, P.O.~Box 2, Green Bank, WV 24944, USA}
\author[0000-0003-2285-0404]{D. R. Madison}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0001-7697-7422]{M. A. McLaughlin}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0003-2397-8290]{S. T. McWilliams}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0002-4307-1322]{C. M. F. Mingarelli}
\affiliation{Department of Physics, University of Connecticut, 196 Auditorium Road, U-3046, Storrs, CT 06269-3046, USA}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Avenue, New York, NY 10010, USA}
\author[0000-0002-3616-5160]{C. Ng}
\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St. George St., Toronto, ON M5S 3H4, Canada}
\author[0000-0002-6709-2566]{D. J. Nice}
\affiliation{Department of Physics, Lafayette College, Easton, PA 18042, USA}
\author[0000-0001-5465-2889]{T. T. Pennucci}
\affiliation{Hungarian Academy of Sciences MTA-ELTE ``Extragalatic Astrophysics Research Group'', Institute of Physics, E\"{o}tv\"{o}s Lor\'{a}nd University, P\'{a}zm\'{a}ny P. s. 1/A, 1117 Budapest, Hungary}
\author[0000-0002-8826-1285]{N. S. Pol}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0001-5799-9714]{S. M. Ransom}
\affiliation{University of Virginia, Department of Astronomy, P.O.~Box 400325, Charlottesville, VA 22904, USA}
\affiliation{National Radio Astronomy Observatory, 520 Edgemont Road, Charlottesville, VA 22903, USA}
\author[0000-0002-5297-5278]{P. S. Ray}
\affiliation{Naval Research Laboratory, Washington DC 20375, USA}
\author[0000-0002-7778-2990]{X. Siemens}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331, USA}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\author[0000-0002-6730-3298]{R. Spiewak}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\affiliation{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, PO Box 218, Hawthorn, VIC 3122, Australia}
\author[0000-0001-9784-8670]{I. H. Stairs}
\affiliation{Department of Physics and Astronomy, University of British Columbia, 6224 Agricultural Road, Vancouver, BC V6T 1Z1, Canada}
\author[0000-0002-1797-3277]{D. R. Stinebring}
\affiliation{Department of Physics and Astronomy, Oberlin College, Oberlin, OH 44074, USA}
\author[0000-0002-7261-594X]{K. Stovall}
\affiliation{National Radio Astronomy Observatory, 1003 Lopezville Rd., Socorro, NM 87801, USA}
\author[0000-0002-1075-3837]{J. K. Swiggum}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee,\\ P.O.~Box 413, Milwaukee, WI 53201, USA}
\author[0000-0002-6428-2620]{R. van Haasteren}
\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\author[0000-0002-6020-9274]{C. A. Witt}
\affiliation{Department of Physics and Astronomy, West Virginia University, P.O.~Box 6315, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Bldg., Morgantown, WV 26505, USA}
\author[0000-0001-5105-4058]{W. W. Zhu}
\affiliation{CAS Key Laboratory of FAST, Chinese Academy of Science, Beijing 100101, China}

% \collaboration{1}{NANOGrav collaboration}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.3 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The regularity of pulsar emissions becomes apparent once we reference the pulses' times of arrivals to the inertial rest frame of the solar system. It follows that errors in the determination of Earth's position with respect to the solar-system barycenter can appear as a time-correlated bias in pulsar-timing residual time series, affecting the searches for low-frequency gravitational waves performed with pulsar timing arrays.
Indeed, recent array datasets yield different gravitational-wave background upper limits and detection statistics when analyzed with different solar-system ephemerides. Crucially, the ephemerides do not generally provide usable error representations.
In this article we describe the motivation, construction, and application of a physical model of solar-system ephemeris uncertainties, which focuses on the degrees of freedom (Jupiter's orbital elements) most relevant to gravitational-wave searches with pulsar timing arrays.
This model, \textsc{BayesEphem}, was used to derive ephemeris-robust results in NANOGrav's 11-yr stochastic-background search, and it provides a foundation for future searches by NANOGrav and other consortia.
The analysis and simulations reported here suggest that ephemeris modeling reduces the gravitational-wave sensitivity of the 11-yr dataset; and that this degeneracy will vanish with improved ephemerides and with the longer pulsar timing datasets that will become available in the near future.
\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{Gravitational waves -- Methods: data analysis -- Pulsars: general -- Ephemerides}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

@@

This article is organized as follows. In Sec.\ \ref{sec:sses} we summarize the SSE production, and we describe the history and stated accuracy of the JPL SSEs adopted in our work; in Sec.\ \ref{sec:systematics} we discuss the TOA delays induced by SSE errors, which are partially absorbed by the timing model, and we identify Jupiter and Saturn's orbits as the drivers behind GW-posterior discrepancies; in Sec.\ \ref{sec:physical} we formulate \textsc{BayesEphem}, and give details about its implementation in our PTA data-analysis software, \textsc{Enterprise} \citep{2019ascl.soft12015E}; in Sec.\ \ref{sec:results} we report on the Bayesian posteriors (for the GW amplitude and for orbital-correction parameters) obtained with \textsc{BayesEphem} for NANOGrav's 11-yr dataset, reproducing and expanding the results of \cite{2018ApJ...859...47A};
in Sec.\ \ref{sec:simulations} we present simulations that probe the reduction in GW sensitivity due to \textsc{BayesEphem};
in Sec.\ \ref{sec:othermodels} we discuss other approaches toward SSE uncertainty modeling;
last, in Sec.\ \ref{sec:conclusion} we offer our brief conclusions.

All computational results presented in this article were obtained with \textsc{Enterprise}  \citep{2019ascl.soft12015E}, used in conjunction with the stochastic sampler \textsc{PTMCMCSampler}\footnote{\href{https://github.com/jellis18/PTMCMCSampler}{https://github.com/jellis18/PTMCMCSampler}} and the pulsar-timing package \textsc{Tempo2} \citep{2012ascl.soft10015H}.
Code that supports all the calculations discussed here and that produces all the figures of this paper is available as a Jupyter notebook at \url{https://github.com/nanograv/11yr_stochastic_analysis}.

@@

These corrections yield delays with the same periods as the corresponding planets, so again they may be observable in PTA datasets for Jupiter and Saturn, but not for Uranus and Neptune.
% MV: say anything about the decay of accuracy past the end of the data?
%
\begin{figure*}[t]
    \centering
    \includegraphics[width=\columnwidth]{figures/roemerdiff.pdf}
    \caption{Differences in R{\o}mer delays computed with DE421/DE430/DE435/DE438 and taking DE436 as reference for 100 simulated pulsars randomly distributed across the sky, plotted after subtracting the best-fitting quadratic and yearly sinusoid. The darker curves with various dashing styles allow a comparison of the same pulsars across the three differences. The plot covers the span of the NANOGrav 11-yr dataset.
    For any given pulsar, these differences are reduced to less than 10 ns after subtracting the best-fitting linear combination of \textsc{BayesEphem} partials, dotted into the pulsar position to obtain R{\o}mer delays.
    }
    \label{fig:Roemer}
\end{figure*}

@@

\input{acknowledge}

@@

%% Similar to \facility{}, there is the optional \software command to allow 
%% authors a place to specify which programs were used during the creation of 
%% the manuscript. Authors should list each code and include either a
%% citation or url to the code inside ()s when available.

%\software{astropy \citep{2013A&A...558A..33A},  
%          Cloudy \citep{2013RMxAA..49..137F}, 
%          SExtractor \citep{1996A&AS..117..393B}
%          }

\bibliography{references}{}
\bibliographystyle{aasjournal}

%% This command is needed to show the entire author+affiliation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges
